---
title: "ST558 Porject 1"
author: "Chien-Lan Hsueh"
institute: Online Statistics Master Program, NCSU
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  html_notebook:
    theme: cerulean
    highlight: haddock
    code_folding: none
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.path = "./images/")
```


## Introduction
[The Open Movie Database (OMDb)](https://www.omdbapi.com/) is a public online database and provides API to use the web service for movie information.

In this project, I will go through my setup including a script to render the rmarkdown, and the tools and libraries I use in this project. Then I will develop a set of helper functions including a wrapper function to make the API calls easy with returned data in a data frame format. After checking all the functions work as expected, I will then use it to obtain some movie information from OMDb web service and perform  some basic EDA including numerical and graphical summaries on the data.

Meanwhile, a [blog](https://chsueh2.github.io/ST558-1st-project.html) is posted on my GitHub pages to share my experience and thinking when doing this project.


## Setup 0: Codes to Render this Markdown Page
This code chunk (don't execute) shows how to use `RMarkdown::render()` to render this rmarkdown document. 
```{r, eval=FALSE}
library(rmarkdown)
library(knitr)

rmarkdown::render(
  input = input, 
  output_format = github_document(html_preview = FALSE), 
  output_file = "README.md"
  output_dir = output_dir
)
```


## Setup 1: Confidential Keys
It's a good practice not to include any login confidential information like passwords or access keys explicitly in codes. Here, we use `keyring` to store the API key.

At the first time to run this notebook, uncomment the following line and execute it. RStudio will ask for your API key and store it. This needs to be done only once.

```{r}
# Uncomment the following line to save the API key
#rstudioapi::askForSecret("OMDb")
```


## Setup 2: (Install and) Load Packages
We will use the following packages in this project:

- `here`: enables easy file referencing and builds file paths in a OS-independent way
- `keyring`: provides OS-independent 'API' to access system's credential store
- `stats`: loads this before loading `tidyverse` to avoid masking some `tidyverse` functions
- `tidyverse`: includes collections of useful packages like `dplyr` (data manipulation), `tidyr` (tidying data),  `ggplots` (creating graphs), etc.
- `rlang`: deals with missing arguments in our helper functions
- `lubridate`: handle date and datetime data type
- `glue`: offers interpreted string literals for easy creation of dynamic messages and labels
- `scales`: formats and labels scales nicely for better visualization
- `httr`: performs HTTP web APIs
- `jsonliet`: parses data stored in JSON format
- `GGally`: plots a very nice scatterplot matrix

In addition, the `pacman` package provides handy tools to manage R packages (install, update, load and unload). We use its `p_laod()` instead of `libarary()` to load the packages listed above. 

```{r, results='hide'}
# load the `pacman` package
# if it's not yet installed, install and load it
if (!require("pacman")) utils::install.packages("pacman", dependencies = TRUE)

# load the required packages whether it's installed or not
pacman::p_load(
	here, keyring,
	stats, 
	tidyverse, rlang,
	lubridate,
	glue, scales,
	httr, jsonlite,
	GGally
)
```

## Helper Functions

The first helper function is a negated version of the built-in `%in%` operator. This makes our codes cleaner and easier to read when we check if an element is absent.
```{r}
# not %in%
'%notin%' <- Negate('%in%')
```

To make the OMDb API easy to use, we define some helper functions and wrap them up in a user-friendly function to perform a query. There are mainly two different API queries:  (1) inquiry by a IMDb ID or a movie title, and (2) inquiry with a movie title search. To perform the queries, we define the following helper functions and an easy-to-use wrapper to do API calls.

- `OMDb_setup_query()`: sets up the query parameters
- `OMDB_query()`: performs an API call and check the results (status)
- `OMDb_parse_movie()`: parse inquiry results of a movie by its IMDb ID or title (with/without other optional parameters)
- `OMDb_parse_search()` parse search results (with/without using other optional parameters)
- `OMDb_movie()`: a wrapper function to do a movie query or search (with or without other optional parameters)

Note: The variable `Year` returned in the query response is character but might not be a single year. In some cases (especially for media type: series), it can be a character string like `2015-2020` or `2015-`. Therefore, `Year` variable will be kept as is. Instead, a new date variable `Year_satrt` will be created to get the beginning year of the span.


### `OMDb_setup_query()`
Create query parameters in a list of key-value pairs:

> Arguments:
>
> - `args`: a list of all other arguments passed in from the wrapper function `OMDb_movie()`
> - `by`: method of API query. Needs to be one of the API supported methods
> - `value`: value to inquery or search for
> - `page`: (optional) specified page of search results
> - `verbose`: a boolean switch to print out the query
>
> Returned Value: query parameters in a list of key-value pairs

```{r}
# a helper function to create query parameters in a list of key-value pairs
OMDb_setup_query <- function(args, by = c("id", "title", "search"), value, page, verbose = T, ...) {
  # verify query method argument `by` to be one of the supported methods
  by <- match.arg(by)
  
  # assign default query value if it's missing in argument (for demo and test purpose)
  if(is_missing(value)){
    value <- c("tt1285016", "Batman", "Batman")[by == c("id", "title", "search")]
  }

  # assign default page value (only needed when using search method)
  if(by == "search") {
    args$page <- ifelse(is_missing(page), "1", as.character(page))
  }
  
  # remove `by` and `value` from the list
  # also remove argument `r` (if exists) because this function only supports JSON
  args[c("by", "value", "r")] <- NULL

  # construct the query as a list of query parameters and the api key
  query0 <- 
    # key-value pair from the query method `by` and query value `value`
    switch(by,
      id = list(i = value),
      title = list(t = value),
      search = list(s = value)) %>% 
    # append the rest of query parameters
    append(args)

  # print out the query
  if(verbose){
    print(glue(
      "Query: {paste(names(query0), query0, sep = '=', collapse = ',')}"
    ))
  }

  # add the api key
  query <- query0 %>% 
    # add the api key
    append(list(apiKey = key_get("RStudio Keyring Secrets", "OMDb")))
  
  return(query)
}
```

### `OMDB_query()`
A wrapper of `httr::Get()` to send a query request, convert and parse the response data, and check the execution status. 

> Arguments:
>
> - `query`: query parameters in a list of key-value pairs
>
> Returned Value: parsed JSON content data from the API response

```{r}
# a low-level helper function to send a query request and check the execution status
OMDB_query <- function(query){
  # send API query request
  rawdata <- GET(url = "http://www.omdbapi.com/", query = query)

  # check if the request is successful
  # if not, stop and print the HTTP status code
  if(!between(rawdata$status_code, 200, 299)) {
    stop(glue(
      "The HTTP GET request is not successful. HTTP status code: {rawdata$status_code}"
    ))
  }
  
  # parse the JSON content data
  parsed <- fromJSON(rawToChar(rawdata$content))
  
  # remove any empty list
  for(i in names(parsed)){
    if(length(parsed[[i]]) == 0) parsed[[i]] <- NA
  }
  
  # failed response
  if(parsed$Response == "False"){
    stop(glue(
      "The HTTR Get request is successful but the respose of the query has a false return ",
      "with an error message from API server: {parsed$Error}"
    ))
  }
  
  return(parsed)
}
```


### `OMDb_parse_movie()`
Parse info of a movie and save in a data frame after data type conversion for numeric data (`Runtim` in minutes, `imdbRating`, `BoxOffice` in dollar amount, etc) and date time data (`Released` date).

> Arguments:
>
> - `parsed`: a parsed JSON content data from a query of a movie
>
> Returned Value: a data frame with the information of the movie

```{r}
# a helper function to parse info of a movie and save in a data frame
OMDb_parse_movie <- function(parsed){
  # convert the data into appropriate data type and save in a data frame
  df <- parsed %>% 
    as_tibble() %>% 
    # convert columns to numeric type
    mutate(across(matches(
      c("Runtime", "Metascore", "imdbRating", "imdbVotes", "BoxOffice", "totalSeasons")),
      ~parse_number(.x, na = c("", "NA", "N/A")))) %>% 
    # convert columns to datetime type
    mutate(across(matches(
      c("Released", "DVD")), 
      ~parse_date(.x, na = c("", "NA", "N/A"), format = "%d %b %Y"))) %>% 
    mutate(Year_start = parse_number(Year))
  
  return(df)
}
```

### `OMDb_parse_search()`
Parse search results and save in a data frame. If only asks for a specified page of the search results, return that page, otherwise, retrieve the remaining pages and return all results.

> Arguments:
>
> - `parsed`: a parsed JSON content data from a query of a movie
> - `query`: query parameters in a list of key-value pairs
> - `args`: a list of all other arguments passed in from the wrapper function `OMDb_movie()`
>
> Returned Value: a data frame with the search results

```{r}
# a helper function to parse search results and save in a data frame
OMDb_parse_search <- function(parsed, query, args){
  df <- parsed$Search %>% as_tibble()
  
  # how many pages of the search results    
  totalPages <- ceiling(as.numeric(parsed$totalResults)/10)
  
  # if there are more than one page and page argument is not specified
  # get the remaining pages
  if (("page" %notin% names(args)) && totalPages > 1){
    for (i in 2:totalPages) {
      query$page <- as.character(i)
      df_nextpage <- OMDB_query(query)$Search %>% as_tibble()
      df <- bind_rows(df, df_nextpage)
    }
  }
  
  # add the total results and sort the data by Type and Year  
  df <- df %>% 
    mutate(
      Year_start = parse_number(Year),
      totalResults = parsed$totalResults
    ) %>% 
    arrange(Title, Type, Year)
}
```

### `OMDb_movie()`
A wrapper function to do a movie query or search (with or without other optional parameters).

> Arguments:
>
> - `by`: method of API query. Needs to be one of the API supported methods
> - `...`: (optional) query parameters including `type`, `y` (year), `page` (page number of search return), `plot`, etc. More info of the API parameters can be found [here](https://www.omdbapi.com/#parameters).
>
> Returned Value: a data frame of what the query returns

```{r}
# a user-friendly wrapper function to perform an API query
OMDb_movie <- function(by = c("id", "title", "search"), ...) {
  # save all arguments as a named list
  args <- as.list(match.call())[-1]
  
  # set up query string
  query <- OMDb_setup_query(args, by, ...)
  parsed <- OMDB_query(query)
  
  # convert parsed data into a data frame based on query method
  if (by == "search"){
    # search returns a list of movies
    df <- OMDb_parse_search(parsed, query, args)
  } else {
    # when using search by id and title,  returns info of the movie
    df <- OMDb_parse_movie(parsed)
  }
  
  return(df)
}
```


## Test and Examples of Using `OMDb_movie()`
In this section, we will test the wrapper function and demonstrate its usages.

### Single-movie queries (by IMDb ID)
When doing a single-movie query with an IMDb ID (using OMDb API's `i` parameter), the API only returns an exact match if there exists, or an false response with an error message if not exact match. The successful returned data can be a movie, a TV series or a game title. Different media types have different variables (data columns) in the returns and the wrapper function should be able to take care of these differences in the background.
```{r}
# Superman
super1 <- OMDb_movie(by = "id", value = "tt0078346") # movie
super1

super2 <- OMDb_movie(by = "id", value = "tt0106057") # series
super2

super3 <- OMDb_movie(by = "id", value = "tt0772202") # game
super3
```

Note: If a query returns a movie, the variable `ratings` is a dateframe with two columns: `Ratings$Source` and `RatingsValue`. In a tibble, this is a list column (a column that stores a dataframe in it). The dataframe printout might look differently depending on the rendering options.



### Single-movie queries (by Title)
When doing a query by a title (using OMDb API's `t` parameter), the API call again returns one result but its behavior is different from queries by IMDb ID in the following ways:

- If there is no exact match, it returns one movie with partially matched title (not clearly how it decides which to pick). 
- If there is only one exact match, it turns that movie.
- If there are more than one exact match, it still returns one movie (again, not clearly how it decides which to pick). We can specify an additional search parameter like `y` (year) to get other matched movies.

```{r}
# Batman
batman1 <- OMDb_movie(by = "title", value = "Batman") # title
batman1

batman2 <- OMDb_movie(by = "title", value = "Batman", y = 1966) # year
batman2
```

We can also specify the parameter `plot` to get different versions of the plot:
```{r}
batman3 <- OMDb_movie(by = "title", value = "Batman", y = 1966, type = "series") # series
batman3
batman4 <- OMDb_movie(by = "title", value = "Batman", y = 1966, type = "series", plot = "full") # series
batman4
```

```{r}
print(glue(
  "Short plot: \n\t{batman3$Plot}\n\n",
  "Full plot: \n\t{batman4$Plot}"
))
```

### Search with a Partially Matched Value
Searches using OMDb API's `s` parameter return a list of media (movies/series/games) as long as they have exact  or partial match to the search value. Without specifying `page` parameter, our helper function will make multiple API calls to retrieve all of the results page by page and return all results in a data frame.

```{r}
batman_list <- OMDb_movie(by = "search", value = "Batman")
batman_list
```

If the `page` parameter is specified, then it will only return the results on the page (10 results on a page):
```{r}
batman_list_p1 <- OMDb_movie(by = "search", value = "Batman", page = 1)
batman_list_p1
batman_list_p2 <- OMDb_movie(by = "search", value = "Batman", page = 2)
batman_list_p2
```

### Tests with Bad Parameters
It's always a good idea to test how our functions handle some bad inputs (query parameters) by users. In this section, we briefly test the following three cases.

#### Bad (Invalid) IMDBb ID
If we have an invalid IMDb ID, we get an error message from the respond. 

Note: I have to turn the evaluation option off or the rendering function won't work.
```{r, eval=FALSE}
bad_id <- OMDb_movie(by = "id", value = "in_valid_ID_tt0772203332")
bad_id

# the error message:
#Error in OMDB_query(query) : 
#The HTTR Get request is successful but the respose of the query has a false return with an error message from API server: Incorrect IMDb ID. 
```

#### Bad (Makeup) Title
It seems this API always tries it best and returns a "partially" matching result even when you use a very weird title to search for:
```{r}
bad_title <- OMDb_movie(by = "title", value = "@!#2@#")
bad_title
```

#### Bad Search Value
Because there is no exact match and are too many partial matched results, the API call return a failure response (although the HTTR request is successful) with an error message (from API server) saying that there are too many results.

Note: I have to turn the evaluation option off or the rendering function won't work.
```{r, eval=FALSE}
bad_list <- OMDb_movie(by = "search", value = "@!#2@#")
bad_list

# the error message:
#Error in OMDB_query(query) : 
#The HTTR Get request is successful but the response of the query has a false return with an error message from API server: Too many results.
```


## EDA
Since we have already done a search to get a list of all Batman-related title, let's take a look at the data.

```{r}
print(glue(
  "The Batman search returns {nrow(batman_list)} results ",
  "in {length(unique(batman_list$Type))} types of media ",
  "including {batman_list$Type %>% unique() %>% toString()} ",
  "from {min(batman_list$Year_start)} to {max(batman_list$Year_start)} ",
  "(planned to be released)."
))

# one-way table
table(batman_list$Type)

# bar plot of media types
batman_list %>% 
  ggplot(aes(x = Type, fill = Type)) +
  geom_bar() +
  labs(
    title = "Comparison of Number of Batman Movies/Series/Game Titles",
    subtitle = glue("From {min(batman_list$Year_start)} to {max(batman_list$Year_start)}"),
    x = "Media Type",
    y = "Count"
  )

# histogram
batman_list %>% ggplot(aes(Year_start, fill = Type)) +
  geom_histogram(position = "dodge2", binwidth = 1) +
  labs(
    title = "Histogram of Batman Movies/Series/Game Titles by Year",
    subtitle = glue("From {min(batman_list$Year_start)} to {max(batman_list$Year_start)}"),
    x = "Released Year",
    y = "Number of Released Titles"
  )
```

I think it will be interesting to look at the movies of DC's another big IP: Superman.

```{r}
superman_list <- OMDb_movie(by = "search", value = "Superman")
superman_list
```

```{r}
print(glue(
  "The SuperBatman search returns {nrow(superman_list)} results ",
  "in {length(unique(superman_list$Type))} types of media ",
  "including {superman_list$Type %>% unique() %>% toString()} ",
  "from {min(superman_list$Year_start)} to {max(superman_list$Year_start)} ",
  "(planned to be released)."
))

# one-way table
table(superman_list$Type)
```
We join these two data sets and compare them in a two-way contingency table:

```{r}
DC_heroes <- bind_rows(
  batman_list %>% mutate(Hero = "Batman"),
  superman_list %>% mutate(Hero = "Superman")
)
DC_heroes

# two-way table: heroes by media type
table(DC_heroes$Type, DC_heroes$Hero)

# faceted scatter plots of counts by year
DC_heroes %>% ggplot(aes(Year_start, color = Hero)) + 
  geom_freqpoly(binwidth = 1) +
  facet_grid(rows = vars(Type), scales = "free") +
  labs(
    title = "Comparison of Number of Movies/Series/Game Titles by Year",
    subtitle = "Batman vs. Superman",
    x = "Released Year",
    y = "Number of Released Titles"
  )
```

From the faceted plot above, we can see that Batman have more titles produced in all three media types. Among the them TV series show strong periodic patterns than the other two media types. 

Next, we will only focus on movies and only those with box office records. For each movie, we will call the API wrapper function `OMDb_movie()` with its IMDb ID to get its information including box office records. To reduce the number of API calls, those movies in `DC_heroes` without poster urls are unlikely to be played in movie theaters and thus will be excluded in our next API queries.

<font color = red>
Note: I have to reduce the number of API calls because it's so easy to reach the API query daily limit. 

In case we need to do a reduced size of study, we canI limit the number of queries by sampling some in the `DC_heroes` using `slice_sample(prop = 0.3)` and in hope that the returned data will have enough Batman and Superman movies with box office records for the later EDA.

To get full search for EDA, set `slice_sample(prop = 1)` or uncomment the line in the following code chunk.
</font>

```{r}
# use API to get movie data for each movies
df_DC_raw <- DC_heroes %>% 
  # only movies with posters
  filter(Type == "movie", Poster != "N/A") %>% 
  #slice_sample(prop = 0.3) %>% 
  select(Hero, imdbID) %>% 
  # convert to a named vector of IMDb ID
  (function(x)(set_names(x$imdbID, x$Hero)))() %>% 
  # map through the IMDb ID to call API for queries of each movie
  map_dfr(~OMDb_movie(by = "id", value = ., verbose = F), .id = "Hero")
```


```{r}
# side-by-side box plots of run time
df_DC_raw %>% 
  ggplot(aes(Hero, Runtime)) +
  geom_boxplot(na.rm = TRUE) +
  geom_jitter(aes(color = Rated), width = 0.1, na.rm = TRUE) +
  labs(
    title = "Runtime of All Batman and Super Movies",
    subtitle = "First Look: Raw data",
    x = "Superhero",
    y = "Runtime [min]"
  )
```

There are `r df_DC_raw$imdbID %>% unique() %>% length()` data rows from the query responses. Roughly speaking, Superman movies have longer run time. It is noticed that many of them are very short and are not rated by [Motion Picture Association  (MPA)](https://en.wikipedia.org/wiki/Motion_Picture_Association_film_rating_system). These short and unrated movies are not likely played in movie theater and have box office records.

In order to make the following analysis more reasonable, I will only consider those with MPA ratings and also remove the repeating data rows due to flattening of the list column `Ratings`. Two new variables are also created:

- `BO_per_min`: Box office dollars per run time in minutes
- `popularity`: Popularity scaled defined as a product of IMDb Rating score and number of votes

```{r}
df_DC <- df_DC_raw %>%
  # remove repeat rows due to the flattening of list column
  select(-Ratings) %>% 
  # filter out non-MPA rated films
  filter(Rated %in% c("G", "PG", "PG-13", "R")) %>% 
  mutate(
    Rated = factor(Rated, levels = c("R", "PG-13", "PG", "G")),
    BO_per_min = BoxOffice/Runtime,
    popularity = imdbRating * imdbVotes
  )

# scatter plot of run time by year
df_DC %>% 
  ggplot(aes(Released, Runtime, color = Hero)) +
  geom_point(na.rm = TRUE) +
  facet_wrap(vars(Rated)) +
  labs(
    title = "Runtime of Batman and Super Movies by Year",
    subtitle = "Films with MPA Ratings",
    x = "Year",
    y = "Runtime [min]"
  )
  
# 2-way table
addmargins(table(df_DC$Hero, df_DC$Rated), c(1, 2))
```

The data set now has `r nrow(df_DC)` movies with `r sum(df_DC$Hero == "Batman")` Batman movies and `r sum(df_DC$Hero == "Superman")` movies. Most of them are rated PG and PG-13. Batman movies have more rated R films probably because its story is darker and more violent. From the graph, it clearly shows that Superman movies on average have longer run time than the Batman movies. In the past decades, there is a trend that the movie run time are getting longer in R and PG-13 rated films. 

Now let's take at look at the box office:
```{r}
# scatter plot of run time by year
df_DC %>% 
  ggplot(aes(Runtime, BoxOffice, color = Hero)) +
  geom_point(na.rm = TRUE) +
  geom_smooth(formula = y ~ x, method = "lm", se = F, na.rm = T) +
  labs(
    title = "Box Office of Batman and Super Movies by Run Time",
    subtitle = "Films with MPA Ratings",
    x = "Runtime [min]",
    y = "Box Office [$]"
  ) +
  scale_y_continuous(labels = label_dollar(suffix = " M", scale = 10^-6))
```

Long movies tend to have bigger box office amount! There is a strong correlation between `BoxOffice` and `Runtime`. They have a large positive correlation coefficient `r cor(df_DC$BoxOffice, df_DC$Runtime, use = "pairwise.complete.obs")`. Below is a side-by-side numeric summaries of the box office amount per run time (in million dollars per minute):
```{r}
# Numeric summaries on box office amount
df_DC %>%
  mutate(
    BoxOffice = BoxOffice / 10^6,
    BO_per_min = BO_per_min / 10^6
  ) %>% 
  drop_na(BoxOffice, BO_per_min) %>% 
  group_by(Hero) %>% 
  summarise(
    P75 = quantile(BO_per_min, probs = 0.75),
    Mean = mean(BO_per_min),
    P25 = quantile(BO_per_min, probs = 0.25),
    Std = sd(BO_per_min)
  )
```

On average, every minute of the Batman movie brings in more box office amount than the Superman movies.

In addition to `Runtime`, it's worthy to take a look at how other variables like `imdbRating` and `imdbVotes` correlate with the box office. A scatterplot matrix comes handy to visualize their relationship.
```{r}
# scatterplot matrix with correlation coefficients
df_DC %>% 
  select(Hero, BoxOffice, Runtime, imdbRating, imdbVotes) %>% 
  drop_na() %>% 
  ggpairs(
    aes(color = Hero, alpha = 0.2),
    axisLabels = "none"
  ) + 
  labs(
    title = "Scatterplot Matrix with Correlation Coefficients"
  )
```

The number of votes on IMDb also has a strong correlation with the box office. 

At last, we will plot a scatter plot to show the popularity score we defined earlier is a strong indicator on the box office.

```{r}
# scatter plot to show the correlation between box office and the popularity score
df_DC %>% 
  ggplot(aes(popularity, BoxOffice, color = Hero)) +
  geom_point(na.rm = TRUE) +
  labs(
    title = "Box Office of Batman and Super Movies vs. Popularity Score",
    subtitle = "Films with MPA Ratings",
    x = "Popularity Score [Voting number times ratings]",
    y = "Box Office [$]"
  ) +
  scale_x_continuous(labels = number_format(scale = 10^-6)) +
  scale_y_continuous(labels = label_dollar(suffix = " M", scale = 10^-6))
```





